### 2024-03-27 tesis_Ashish/Chapter4/chapter4.tex
update 28 Mar

For studying the performance of PCC luminometer during 2018 data taking, we have divided CMS luminosity data taking into seven periods: A, B, C, D1, D2, D3, and D4, as depicted in % 2018 CMS luminosity data taking is divided into seven periods namely A, B, C, D1, D2, D3 and D4
Fig. \ref{fig:period_bound}. Run ranges for each period are shown in Table \ref{tab:period_run_ranges}.
During period B, specifically corresponding to Fill 6868, van der Meer (vdM) scans were conducted for calibrating the PCC luminometer.
%vdM data is shown during period B corresponding to Fill 6868.

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/period_boundary.png}
\caption[2018 CMS Luminosity Data Period]{%
   2018 luminosity data taking periods showing data taking period boundaries and vdM calibration data (1 July in period B)  \cite{CERNLumiPublicResults}.
}
\label{fig:period_bound}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/good_bad_modules_1.png}
\caption[Good/Bad Module Weights]{%
   Change in normalized module weight with time. Modules showing significant change in module weight (shown in right figure) over time having large RMS/mean values are used to make module veto list.
}
\label{fig:goodbadmodules}
\end{figure}

@@

Module veto list consisting of bad modules is first created for period B, vdM Fill and then for other periods A, C, D1, D2, D3 and D4 as follows,

Run 2018B and vdM Fill:
\begin{itemize}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/cut_selection_1.png}
\caption[First Iteration For Outlier Module Removal]{%
   A loose selection of 7\% is applied to remove bad modules. Appropriate selections are applied in each iterative step to remove other bad modules.
}
\label{fig:outliermodules}
\end{figure}

@@

    \caption[Second and Final Iteration For Outlier Module Removal]{%
        Left: Second iteration on RMS/mean of 3\% is applied to remove more bad modules. 
        Right: Final iteration on RMS/mean of 2\% is applied to obtain module veto list for each period.
    }
    \label{fig:sec_it_cut}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.7\textwidth]{ashish_thesis/mod_weight_comp_2.png}
\caption[Module Weights Change Removal]{%
   Projection of change in module weights between period A and B. Modules showing more than 3 sigma change in module weights are added to the module veto list.
}
\label{fig:mod_w_com_1}
\end{figure}

@@

\begin{table}[!ht]
\centering
%\topcaption{%
 %PCC/ HFOC luminosity mean and rms values for new per period vetolist showing variation in mean values across period.
%}
\begin{tabular}{ccccc}
    \textbf{Period} & \textbf{mean} & \textbf{rms} &  \textbf{rms/mean} \\ \hline
    A & 1.006 & 0.004183 & 0.004158 \\
    B & 0.9981  & 0.003133 & 0.0031389 \\
    C & 0.9989  & 0.003009 & 0.003012\\
    D1 & 1.003  & 0.002441 & 0.002433 \\
    D2 & 1.001  & 0.0029 & 0.00289 \\
   D3  & 1  & 0.002115 & 0.002115 \\
   D4  & 1.002  & 0.003673 & 0.003666 \\ 
\end{tabular}
\caption[Mean value change for PCC/HFOC across period.]{%                                                                                                                                       
 PCC/ HFOC luminosity mean and rms values for  per period veto list showing variation in mean values across period.
}
\label{tab:pccvis_diffveto}
\end{table}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/before_after_vdm_stability_1.png}
\caption[PCC vdM Stability]{%                                                                                                                                                                      
   Luminosity fraction per layer or disk (stability plots) before (no veto) and after applying the final module veto with 2\% RMS/mean selection for the vdM fill data.
}
\label{fig:b_a_stability_vdm}
\end{figure}

@@

\begin{table}
  \centering
  \caption{Run Period and Good Modules}
\begin{tabular}{cccc}
\textbf{Run period (Old Veto)} & \textbf{Good modules} & \textbf{Run period (New Veto)} & \textbf{Good modules} \\
\hline
A & 787 & A & 580 \\
B & 787 & B & 1054 \\
C & 787 & C & 780 \\
D early & 787 & D1 & 687 \\
 & & D2 & 672 \\
 &  & D3 (323363-323699)  &  775\\
D late & 469 & D3 (323700-324420) & 775 \\
&  & D4 & 824\\
\end{tabular}
%\caption{Run Period and Good Modules}
\label{tab:runperiod_oldnew}
\end{table}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=\textwidth]{ashish_thesis/2percentcommonveto_pixellayerdisk_1.png}
\caption[PCC Stability For Physics Runs]{Left: Luminosity fraction of pixel detector for various layer and disks as a function of lumi section for a module selection excluding only the BPIX Layer 0 for the full 2018 zero-bias dataset. Right: Stability profiles of pixel detector layer and disk for final module veto list.}
\label{fig:stabprof_4333}
\end{figure}

@@

\begin{figure}[!htp]
    \centering
    \includegraphics[width=1\textwidth]{ashish_thesis/vdm_program_2018.png}
    \caption[2018 vdM Program]{Beam position along x and y is plotted as a function of time showing different types of scans. #3 and #10 are vdM scans, #4 and #9 are offset scans, #5, #7 and #8 are beam imaging scans, #6 is emittance scan \cite{pas_18}.}
    \label{fig:vdm_prog_2018}
\end{figure}

The vdM data (LHC fill 6868) is processed using the final module selection. The background estimation for the PCC rate is done with two super-separation scans SS1 and SS2. The mean and error are obtained from the $y$ projection of PCC rates per NB4 plotted as a function of time as shown in Fig. \ref{fig:sigmavis_ss_backg}. The results for five different BCIDs are shown in Table \ref{tab:vdm:SS1_SS2}, using the reprocessed PCC data for Fill 6868. The overall background correction applied to raw PCC rate is $0.02757\pm0.01987$, which is calculated by taking the average of the mean values of SS1 and SS2.

@@

\begin{table}
  \begin{center}
    \caption[2018 PCC Background]{Mean for the background estimated with the SS1 and SS2 data, separately for all five BCIDs and averaged.}
    \begin{tabular}{ccccc}
    \textbf{BCID}   & \textbf{Mean (SS1)} & \textbf{Mean (SS2)} \\ \hline
      265     &  0.02902    &  0.02743    \\
        865  &    0.02572  &     0.0281  \\
       1780    &  0.02862   &     0.0286  \\
       2192   &  0.02729  &     0.02323  \\
       3380  &  0.02882  &    0.02896   \\
       Average & 0.02789 & 0.02726  \\
      \end{tabular}
    %\caption[Background in PCC rate]{Mean for the background estimated with the SS1 and SS2 data, separately for all five BCIDs and averaged.}
    \label{tab:vdm:SS1_SS2}
  \end{center}
\end{table}

@@

\begin{figure}[!htp]
    \centering
    \includegraphics[width=1\textwidth]{ashish_thesis/vdM_fit_cveto_1.png}
    \caption[PCC Rate Fit]{Rates and the resulting fitted Poly2G scan curves as a function of the beam separation for a single bunch (BCID 265) as recorded by PCC for vdM1 scan in the x (left) and y direction (right). Background subtraction and the beam corrections have been applied to the raw data before the fit.}
    \label{fig:fitquality}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.7\textwidth]{ashish_thesis/sigma_vis_btob_var_1.png}
\caption[$\sigma_{vis}$ Bunch Variation]{%                                                                                                                                                       
 Bunch to bunch variation of PCC visible cross section.
}
\label{fig:sigmavis_btob_variation}
\end{figure}

@@

The LHC operates in cycles known as "fills." A fill is a period during which the LHC is filled with protons and provides collisions for its experiments. The luminosity profile of a fill is a graphical representation of the instantaneous luminosity as a function of time for the entire duration of that fill. Before the actual data taking begins, there is a phase of beam injection, acceleration, and stabilization. During this period, no collisions are taking place, so the luminosity is zero. Once the beams are accelerated to their maximum energy and brought into collision, there is a rapid rise in luminosity. This is the point where the luminosity reaches its highest value, known as the peak luminosity. After the peak, the luminosity starts to decrease. This period, known as "stable beams", is when most of the data is collected. The reason for the decrease in luminosity over time is the reduction in the number of protons in the beams due to collisions.
%As the protons in the beams collide, the number of protons, and hence the luminosity, diminishes.
The rate of this decrease can be influenced by various factors, including beam dynamics, losses, and the efficiency of the collider's operation. The fill concludes when the number of protons is too low to provide useful data, or if there's a technical issue, leading to a beam dump. At this point, the luminosity drops to zero. The total duration of a fill can vary. Some fills last only a few hours, while others can extend for over a day, depending on operational efficiency and the goals for that specific fill. Fill profile for Fill 6961 during 2018 data taking is shown in Fig. \ref{fig:Fill6961} with

@@

\begin{figure}[!htp]
    \centering
    \includegraphics[width=0.9\textwidth]{ashish_thesis/Fill_profile_6961_1.png}
    \caption[Fill 6961 Profile]{Typical lumi profile for LHC Fill during 2018 data taking period for Fill 6961.}
    \label{fig:Fill6961}
\end{figure}

@@

Fill 9036, shown in \ref{fig:af_fit40} and encompassing 900 colliding bunches, serves as the foundation for estimating the Type 2 afterglow parameters. A fit is performed to one wagon train, then two, and finally three, refining the parameters for Type 2 afterglow and minimizing the residuals of both afterglow types. Fit plots for single train is shown in Fig. \ref{fig:af_fit99}. The values of type 2 afterglow parameter amplitude (B) and decay width (C) for different wagon trains are shown in Table Fig. \ref{fig:af_fit4000}.

@@

\newpage
\begin{figure}[H]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/fill_7036_pattern_2.png}
\caption[7036 Fill Patern]{%                                                                                                                                                                                         
  Fill 7036 pattern (900 b fill).
}
\label{fig:af_fit40}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics[width=0.75\textwidth]{ashish_thesis/2018_af_fit_1.png} 
  %\hfill % or \hspace{5mm} for a specific horizontal space                                                                                                  
  %\begin{subfigure}[b]{0.49\textwidth}
   % \includegraphics[width=\textwidth]{ashish_thesis/2018_af_fit_residuals.png}                                                                                     
  %\end{subfigure}
  \caption[Afterglow Effect Fit]{Fit to single train to estimate afterglow parameters.}
  \label{fig:af_fit99}
\end{figure}

@@

Instantaneous luminosity is proportional to the rate of collisions at any given moment in time. It is a snapshot of the experiment's performance and is represented as \( L_{\text{inst}} \). The unit of instantaneous luminosity is often Hertz per micro barn $(Hz/ub)$. Integrated luminosity is the accumulation of instantaneous luminosity over a specific time period. Mathematically, it is defined as:
\[
\text{Integrated Luminosity} = \int (\text{Instantaneous Luminosity}) \, dt
\]
The unit of integrated luminosity is typically inverse femtobarns (\( \text{fb}^{-1} \)). Integrated luminosity can be understood as the sum of all instantaneous luminosities over the duration of the experiment, enhancing the statistical significance of the data collected. It serves as a measure of the total number of events recorded by the experiment, crucial for robust statistical analysis. When plotted as a function of lumi sections, PCC instantaneous and integrated luminosity offer distinct yet complementary insights into the experiment's performance. In plots that feature luminosity as a function of these lumi sections, the x-axis enumerates the lumi sections while the y-axis presents the luminosity values.  For instantaneous luminosity, discrete data points are plotted for each lumi section, providing immediate, lumi section specific measurements of the collision rate as shown in Fig. \ref{fig:PCC_inst_2018}. These points can fluctuate based on a variety of factors, such as beam conditions or adjustments to the accelerator. On the other hand, integrated luminosity is represented by a continuously ascending curve. %Each point on this curve accumulates the instantaneous luminosity values up to that specific lumi section. %The integrated luminosity for entire 2018 is calculated to be 60.47 fb^{-1}.

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/PCClumivslumisection_2018_1.png}
\caption[2018 PCC Inst. Luminosity]{%                                                                                                                                                                   
 2018 PCC instantaneous luminosity as a function of lumi section.
}
\label{fig:PCC_inst_2018}
\end{figure}

@@

\item Orbit Drift: Orbit drift in particle accelerators refers to the slight deviation or shift in the trajectory of circulating particle beams from their intended path. This drift can be caused by a myriad of factors, including imperfections in the magnetic fields guiding the beams, external disturbances, and even minute ground movements or temperature fluctuations affecting the accelerator components. The precision with which beams are aligned in colliders like the LHC is crucial for maximizing the number of particle collisions, and hence the luminosity. When an orbit drift occurs, the overlap of the colliding beams may become suboptimal, leading to a reduced instantaneous luminosity. This reduction can influence the effectiveness of the experiments conducted, potentially introducing uncertainties in the luminosity measurement \cite{CERNOribLumi}.

@@

\item  X-Y non-factorization: The luminosity calibration often relies on methods that factorize the beam overlap integral into two separate integrals in the horizontal (X) and vertical (Y) directions. When this factorization is not valid or precise, the XY non-factorization comes into play. It essentially refers to the systematic uncertainty arising from the inability to perfectly separate or factorize the beam profiles in the horizontal and vertical dimensions. A variety of reasons can lead to XY non-factorization. One of the main causes is the complex shape and dynamics of the colliding beams. While in an ideal scenario, the beams might be perfectly Gaussian in shape and independent in X and Y dimensions, real-world effects such as beam-beam interactions, intrabeam scattering, or external perturbations can introduce correlations between the two dimensions. This means that the actual overlap of the beams cannot be simply represented as a product of their profiles in X and Y. The luminosity is a direct measure of how many collisions are happening per unit of time and area, and it depends on the overlap of the colliding beams. If the beams' profiles in the X and Y dimensions are not factorized correctly, it can lead to an inaccurate representation of their overlap, thus introducing a systematic uncertainty in the luminosity measurement \cite{CERNLUMPOG}.

@@

\item Beam-Beam Deflection: refers to the change in trajectory of the individual particles in the beam as they pass near each other in the interaction point, due to their electromagnetic interaction \cite{Herr:1982430}. In a perfect scenario, each beam of protons would stay focused and keep their trajectory constant throughout the process. However, when two beams cross each other, the electromagnetic fields of the protons in one beam can exert forces on the protons in the other beam, causing them to deviate from their initial trajectory \cite{babaev2023impact}.

@@

\item Dynamic $\beta$: The beta function ($\beta$*) in the context of particle accelerators like the LHC is a measure of the beam size at the collision point \cite{CERNVdMOMC}, and it plays a crucial role in determining the luminosity of the collisions. Lower values of $\beta$* lead to smaller beam sizes at the interaction point and thus higher luminosity, as the probability of proton-proton collisions is increased. However, $\beta$* is not necessarily constant during the operation of the LHC. It can vary due to factors such as changes in the accelerator's operating conditions, beam-beam interactions. %This variability, often referred to as "dynamic beta", can contribute to uncertainties in the luminosity measurement.
 When $\beta$* changes, it can alter the size of the proton beams at the interaction point. As the luminosity is inversely proportional to the square of the beam size, a small change in $\beta$* can lead to a significant change in luminosity. If these changes in $\beta$* are not accurately tracked and accounted for, they can introduce uncertainties in the luminosity measurement. The challenge is that $\beta$* is not directly observable, so its value and any changes in it have to be inferred from other measurements, such as the beam size and position, the orbit drift.
 %and so on. These measurements have their own uncertainties, which can further contribute to the uncertainty in β* and thus the luminosity.
  %Refers to the time-varying beta function of a particle beam, which describes the rate at which the beam diverges or converges as it propagates. This effect can be important for beam stability and the control of beam emittance
 % \cite{CERNVdMOMC}.

\item Beam Current Calibration:  Beam current calibration ensures that the \( N \) value (number of particles per bunch) used in the luminosity equation is accurate. If the calibration is off, the derived value of \( N \) from the instruments like beam current transformers will be inaccurate, leading to errors in the calculated luminosity $\( L_{inst} \)$. As the luminosity dependence on number of particles per bunch is \( N^2 \), even small errors in beam current calibration can significantly affect the accuracy of $\( L_{inst} \)$ \cite{CERNLumiDays}.

>>>>

For studying the performance of PCC luminometer during 2018 data taking, we have divided CMS luminosity data taking into seven periods: A, B, C, D1, D2, D3, and D4, as depicted in % 2018 CMS luminosity data taking is divided into seven periods namely A, B, C, D1, D2, D3 and D4
Fig. \ref{fig:period_bound}. Run ranges for each period are shown in Table \ref{tab:period_run_ranges}.
During period B, specifically corresponding to fill 6868, van der Meer (vdM) scans were conducted for calibrating the PCC luminometer.
%vdM data is shown during period B corresponding to Fill 6868.

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/period_boundary.png}
\caption[2018 CMS luminosity data period]{%
   2018 luminosity data taking periods showing data taking period boundaries and vdM calibration data (1 July in period B)  \cite{CERNLumiPublicResults}.
}
\label{fig:period_bound}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/good_bad_modules_1.png}
\caption[Good/Bad module weights]{%
   Change in normalized module weight with time. Modules showing significant change in module weight (shown in right figure) over time having large RMS/mean values are used to make module veto list.
}
\label{fig:goodbadmodules}
\end{figure}

@@

Module veto list consisting of bad modules is first created for period B, vdM fill and then for other periods A, C, D1, D2, D3 and D4 as follows,

Run 2018B and vdM fill:
\begin{itemize}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/cut_selection_1.png}
\caption[First iteration for outlier modules]{%
   A loose selection of 7\% is applied to remove bad modules. Appropriate selections are applied in each iterative step to remove other bad modules.
}
\label{fig:outliermodules}
\end{figure}

@@

    \caption[Second and final iteration for outlier modules]{%
        Left: Second iteration on RMS/mean of 3\% is applied to remove more bad modules. 
        Right: Final iteration on RMS/mean of 2\% is applied to obtain module veto list for each period.
    }
    \label{fig:sec_it_cut}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.7\textwidth]{ashish_thesis/mod_weight_comp_2.png}
\caption[Module weights change removal]{%
   Projection of change in module weights between period A and B. Modules showing more than 3 sigma change in module weights are added to the module veto list.
}
\label{fig:mod_w_com_1}
\end{figure}

@@

\begin{table}[!ht]
\centering
%\topcaption{%
 %PCC/ HFOC luminosity mean and rms values for new per period vetolist showing variation in mean values across period.
%}
\begin{tabular}{ccccc}
    \textbf{Period} & \textbf{mean} & \textbf{rms} &  \textbf{rms/mean} \\ \hline
    A & 1.006 & 0.004183 & 0.004158 \\
    B & 0.9981  & 0.003133 & 0.0031389 \\
    C & 0.9989  & 0.003009 & 0.003012\\
    D1 & 1.003  & 0.002441 & 0.002433 \\
    D2 & 1.001  & 0.0029 & 0.00289 \\
   D3  & 1  & 0.002115 & 0.002115 \\
   D4  & 1.002  & 0.003673 & 0.003666 \\ 
\end{tabular}
\caption[Mean value change for PCC/HFOC across period]{%                                                                                                                                       
 PCC/ HFOC luminosity mean and rms values for  per period veto list showing variation in mean values across period.
}
\label{tab:pccvis_diffveto}
\end{table}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/before_after_vdm_stability_1.png}
\caption[PCC vdM stability]{%                                                                                                                                                                      
   Luminosity fraction per layer or disk (stability plots) before (no veto) and after applying the final module veto with 2\% RMS/mean selection for the vdM fill 6868 data.
}
\label{fig:b_a_stability_vdm}
\end{figure}

@@

\begin{table}
  \centering
  \caption{Run period and good modules}
\begin{tabular}{cccc}
\textbf{Run period (Old Veto)} & \textbf{Good modules} & \textbf{Run period (New Veto)} & \textbf{Good modules} \\
\hline
A & 787 & A & 580 \\
B & 787 & B & 1054 \\
C & 787 & C & 780 \\
D early & 787 & D1 & 687 \\
 & & D2 & 672 \\
 &  & D3 (323363-323699)  &  775\\
D late & 469 & D3 (323700-324420) & 775 \\
&  & D4 & 824\\
\end{tabular}
%\caption{Run Period and Good Modules}
\label{tab:runperiod_oldnew}
\end{table}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=\textwidth]{ashish_thesis/2percentcommonveto_pixellayerdisk_1.png}
\caption[PCC stability for physics runs]{Left: Luminosity fraction of pixel detector for various layer and disks as a function of lumi section for a module selection excluding only the BPIX Layer 0 for the full 2018 zero-bias dataset. Right: Stability profiles of pixel detector layer and disk for final module veto list.}
\label{fig:stabprof_4333}
\end{figure}

@@

\begin{figure}[!htp]
    \centering
    \includegraphics[width=1\textwidth]{ashish_thesis/vdm_program_2018.png}
    \caption[2018 vdM program]{Beam position along x and y is plotted as a function of time showing different types of scans. #3 and #10 are vdM scans, #4 and #9 are offset scans, #5, #7 and #8 are beam imaging scans, #6 is emittance scan \cite{pas_18}.}
    \label{fig:vdm_prog_2018}
\end{figure}

The vdM data (LHC fill 6868) is processed using the final module selection. The background estimation for the PCC rate is done with two super-separation scans SS1 and SS2. The mean and error are obtained from the $y$ projection of PCC rates per NB4 plotted as a function of time as shown in Fig. \ref{fig:sigmavis_ss_backg}. The results for five different BCIDs are shown in Table \ref{tab:vdm:SS1_SS2}, using the reprocessed PCC data for fill 6868. The overall background correction applied to raw PCC rate is $0.02757\pm0.01987$, which is calculated by taking the average of the mean values of SS1 and SS2.

@@

\begin{table}
  \begin{center}
    \caption[2018 PCC background]{Mean for the background estimated with the SS1 and SS2 data, separately for all five BCIDs and averaged.}
    \begin{tabular}{ccccc}
    \textbf{BCID}   & \textbf{Mean (SS1)} & \textbf{Mean (SS2)} \\ \hline
      265     &  0.02902    &  0.02743    \\
        865  &    0.02572  &     0.0281  \\
       1780    &  0.02862   &     0.0286  \\
       2192   &  0.02729  &     0.02323  \\
       3380  &  0.02882  &    0.02896   \\
       Average & 0.02789 & 0.02726  \\
      \end{tabular}
    %\caption[Background in PCC rate]{Mean for the background estimated with the SS1 and SS2 data, separately for all five BCIDs and averaged.}
    \label{tab:vdm:SS1_SS2}
  \end{center}
\end{table}

@@

\begin{figure}[!htp]
    \centering
    \includegraphics[width=1\textwidth]{ashish_thesis/vdM_fit_cveto_1.png}
    \caption[PCC rate fit]{Rates and the resulting fitted Poly2G scan curves as a function of the beam separation for a single bunch (BCID 265) as recorded by PCC for vdM1 scan in the x (left) and y direction (right). Background subtraction and the beam corrections have been applied to the raw data before the fit. Residual is defined as (Data-fit)/fit}
    \label{fig:fitquality}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.7\textwidth]{ashish_thesis/sigma_vis_btob_var_1.png}
\caption[$\sigma_{vis}$ bunch variation]{%                                                                                                                                                       
 Bunch to bunch variation of PCC visible cross section.
}
\label{fig:sigmavis_btob_variation}
\end{figure}

@@

The LHC operates in cycles known as "fills." A fill is a period during which the LHC is filled with protons and provides collisions for its experiments. The luminosity profile of a fill is a graphical representation of the instantaneous luminosity as a function of time for the entire duration of that fill. Before the actual data taking begins, there is a phase of beam injection, acceleration, and stabilization. During this period, no collisions are taking place, so the luminosity is zero. Once the beams are accelerated to their maximum energy and brought into collision, there is a rapid rise in luminosity. This is the point where the luminosity reaches its highest value, known as the peak luminosity. After the peak, the luminosity starts to decrease. This period, known as "stable beams", is when most of the data is collected. The reason for the decrease in luminosity over time is the reduction in the number of protons in the beams due to collisions.
%As the protons in the beams collide, the number of protons, and hence the luminosity, diminishes.
The rate of this decrease can be influenced by various factors, including beam dynamics, losses, and the efficiency of the collider's operation. The fill concludes when the number of protons is too low to provide useful data, or if there's a technical issue, leading to a beam dump. At this point, the luminosity drops to zero. The total duration of a fill can vary. Some fills last only a few hours, while others can extend for over a day, depending on operational efficiency and the goals for that specific fill. fill profile for fill 6961 during 2018 data taking is shown in Fig. \ref{fig:Fill6961} with

@@

\begin{figure}[!htp]
    \centering
    \includegraphics[width=0.9\textwidth]{ashish_thesis/Fill_profile_6961_1.png}
    \caption[fill 6961 profile]{Typical lumi profile for LHC fill during 2018 data taking period for fill 6961.}
    \label{fig:Fill6961}
\end{figure}

@@

fill 9036, shown in \ref{fig:af_fit40} and encompassing 900 colliding bunches, serves as the foundation for estimating the Type 2 afterglow parameters. A fit is performed to one wagon train, then two, and finally three, refining the parameters for Type 2 afterglow and minimizing the residuals of both afterglow types. Fit plots for single train is shown in Fig. \ref{fig:af_fit99}. The values of type 2 afterglow parameter amplitude (B) and decay width (C) for different wagon trains are shown in Table Fig. \ref{fig:af_fit4000}.

@@

\newpage
\begin{figure}[H]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/fill_7036_pattern_2.png}
\caption[7036 fill patern]{%                                                                                                                                                                                         
  fill 7036 pattern (900 b fill).
}
\label{fig:af_fit40}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics[width=0.75\textwidth]{ashish_thesis/2018_af_fit_1.png} 
  %\hfill % or \hspace{5mm} for a specific horizontal space                                                                                                  
  %\begin{subfigure}[b]{0.49\textwidth}
   % \includegraphics[width=\textwidth]{ashish_thesis/2018_af_fit_residuals.png}                                                                                     
  %\end{subfigure}
  \caption[Afterglow effect fit]{Fit to single train to estimate afterglow parameters.}
  \label{fig:af_fit99}
\end{figure}

@@

Instantaneous luminosity is proportional to the rate of collisions at any given moment in time. It is a snapshot of the experiment performance and is represented as \( L_{\text{inst}} \). The unit of instantaneous luminosity is often Hertz per micro barn $(Hz/ub)$. Integrated luminosity is the accumulation of instantaneous luminosity over a specific time period. Mathematically, it is defined as:
\[
\text{Integrated Luminosity} = \int (\text{Instantaneous Luminosity}) \, dt
\]
The unit of integrated luminosity is typically inverse femtobarns (\( \text{fb}^{-1} \)). Integrated luminosity can be understood as the sum of all instantaneous luminosities over the duration of the experiment, enhancing the statistical significance of the data collected. It serves as a measure of the total number of events recorded by the experiment, crucial for robust statistical analysis. When plotted as a function of lumi sections, PCC instantaneous and integrated luminosity offer distinct yet complementary insights into the experiment's performance. In plots that feature luminosity as a function of these lumi sections, the x-axis enumerates the lumi sections while the y-axis presents the luminosity values.  For instantaneous luminosity, discrete data points are plotted for each lumi section, providing immediate, lumi section specific measurements of the collision rate as shown in Fig. \ref{fig:PCC_inst_2018}. These points can fluctuate based on a variety of factors, such as beam conditions or adjustments to the accelerator. On the other hand, integrated luminosity is represented by a continuously ascending curve. %Each point on this curve accumulates the instantaneous luminosity values up to that specific lumi section. %The integrated luminosity for entire 2018 is calculated to be 60.47 fb^{-1}.

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/PCClumivslumisection_2018_1.png}
\caption[2018 PCC inst. luminosity]{%                                                                                                                                                                   
 2018 PCC instantaneous luminosity as a function of lumi section.
}
\label{fig:PCC_inst_2018}
\end{figure}

@@

\item Orbit Drift: Orbit drift in particle accelerators refers to the slight deviation or shift in the trajectory of circulating particle beams from their intended path. This drift can be caused by a myriad of factors, including imperfections in the magnetic fields guiding the beams, external disturbances, and even minute ground movements or temperature fluctuations affecting the accelerator components. The precision with which beams are aligned in colliders like the LHC is crucial for maximizing the number of particle collisions, and hence the luminosity. When an orbit drift occurs, the overlap of the colliding beams may become suboptimal, leading to a reduced instantaneous luminosity. This reduction can influence the effectiveness of the experiments conducted, potentially introducing uncertainties in the luminosity measurement \cite{CERNOribLumi}. To calculate the systematic uncertainty due to orbit drift in luminosity measurement, the process begins with measuring the particle beam positions during vdM scans using systems like DOROS BPM and arc BPMs. Corrections are applied based on linear interpolations of these positions to account for any drift from the intended paths, crucial for maintaining optimal beam overlaps and accurate collision counts. The impact of these corrections on the visible cross section is then assessed. Systematic uncertainty is estimated by comparing corrected visible cross section values from both measurement systems, with an additional peak position correction applied for discrepancies at the head-on position. This comprehensive approach consolidates the uncertainties from orbit drift and peak position corrections into a single value.

%Orbit drift impacts beam path alignment, crucial for maximizing collision rates and, consequently, luminosity, use systems like DOROS BPM and arc BPMs to measure beam positions during vdM scans, detecting any drift, correct for orbit drift by linearly interpolating measured beam positions around each VdM scan, adjusting nominal beam separations to reflect actual positions, evaluate how these corrections affect the visible cross section, compare corrected visible cross section values from DOROS and arc BPM measurements to estimate uncertainty. The systematic uncertainty encompasses both orbit drift and peak position corrections, ensuring accuracy in luminosity measurements essential for particle physics experiments.

@@

\item  X-Y non-factorization: The luminosity calibration often relies on methods that factorize the beam overlap integral into two separate integrals in the horizontal (X) and vertical (Y) directions. When this factorization is not valid or precise, the XY non-factorization comes into play. It essentially refers to the systematic uncertainty arising from the inability to perfectly separate or factorize the beam profiles in the horizontal and vertical dimensions. A variety of reasons can lead to XY non-factorization. One of the main causes is the complex shape and dynamics of the colliding beams. While in an ideal scenario, the beams might be perfectly Gaussian in shape and independent in X and Y dimensions, real-world effects such as beam-beam interactions, intrabeam scattering, or external perturbations can introduce correlations between the two dimensions. This means that the actual overlap of the beams cannot be simply represented as a product of their profiles in X and Y. The luminosity is a direct measure of how many collisions are happening per unit of time and area, and it depends on the overlap of the colliding beams. If the beams' profiles in the X and Y dimensions are not factorized correctly, it can lead to an inaccurate representation of their overlap, thus introducing a systematic uncertainty in the luminosity measurement \cite{CERNLUMPOG}. To determine the systematic uncertainty due to X-Y non-factorization in luminosity, a series of specialized beam imaging scans are conducted to capture the actual profiles and overlaps of the colliding beams, challenging the common assumption that beam densities in the X and Y dimensions can be independently factorized. Data from these scans are used to model the beam shapes, incorporating potential X-Y correlations using various mathematical functions. Simulated vdM scans based on these models then predict the expected luminosity, accounting for the non-factorizable nature of the beams. By comparing these predictions with actual vdM scan results which assume factorizable beam profiles a correction factor is estimated to adjust the luminosity. The systematic uncertainty related to X-Y non-factorization is finally determined by evaluating the correction factor variability and the model fit quality, ensuring the adjustment accurately reflects the true beam interactions and their impact on the measured luminosity.

@@

\item Beam-Beam Deflection: refers to the change in trajectory of the individual particles in the beam as they pass near each other in the interaction point, due to their electromagnetic interaction \cite{Herr:1982430}. In a perfect scenario, each beam of protons would stay focused and keep their trajectory constant throughout the process. However, when two beams cross each other, the electromagnetic fields of the protons in one beam can exert forces on the protons in the other beam, causing them to deviate from their initial trajectory \cite{babaev2023impact}. Systematic uncertainty due to beam-beam deflection is computed by modeling the electromagnetic interactions that cause proton beams to deflect when collidingaffecting luminosity. Sophisticated simulations predict the deflection impact, allowing for the derivation of a correction factor to adjust luminosity. This uncertainty is quantified by evaluating the correction factor variability and confidence across various conditions.

@@

\item Dynamic $\beta$: The beta function ($\beta$*) in the context of particle accelerators like the LHC is a measure of the beam size at the collision point \cite{CERNVdMOMC}, and it plays a crucial role in determining the luminosity of the collisions. Lower values of $\beta$* lead to smaller beam sizes at the interaction point and thus higher luminosity, as the probability of proton-proton collisions is increased. However, $\beta$* is not necessarily constant during the operation of the LHC. It can vary due to factors such as changes in the accelerator's operating conditions, beam-beam interactions. %This variability, often referred to as "dynamic beta", can contribute to uncertainties in the luminosity measurement.
 When $\beta$* changes, it can alter the size of the proton beams at the interaction point. As the luminosity is inversely proportional to the square of the beam size, a small change in $\beta$* can lead to a significant change in luminosity. If these changes in $\beta$* are not accurately tracked and accounted for, they can introduce uncertainties in the luminosity measurement. The challenge is that $\beta$* is not directly observable, so its value and any changes in it have to be inferred from other measurements, such as the beam size and position, the orbit drift.
 %and so on. These measurements have their own uncertainties, which can further contribute to the uncertainty in β* and thus the luminosity.
  %Refers to the time-varying beta function of a particle beam, which describes the rate at which the beam diverges or converges as it propagates. This effect can be important for beam stability and the control of beam emittance
 % \cite{CERNVdMOMC}.
The systematic uncertainty due to dynamic $\beta$ effects is quantified based on beam transport simulations that incorporate both the defocusing effects of beam-beam interactions and the scaling of these effects to the actual beam energies, $\beta$ settings, and measured beam widths. The correction and its uncertainty are derived from the modeling of how each beam, under the influence of the other, experiences changes in its focusing properties, thereby affecting the overlap and, consequently, the luminosity. The correction is essential to account for the fact that the presence of one beam influences the focusing properties of the other, leading to adjustments in the measured beam sizes and the calculated luminosity.
